Easy as Pi: A Network Coding Raspberry Pi Testbed

### 2016-06-04 introduction.tex
Starting with the results
Update Raspi acronym

Our work is organized as follows: Section \ref{sec:testbed} introduces
the testbed system, its setup, connectivity, files configuration and
compilation of the Kodo library. Section \ref{sec:schemes} defines the
coding schemes employed in our study. Section \ref{sec:topologies} some
configurable topologies in the tested. Later, in Section \ref{sec:metrics}
we describe the considered metrics for performance comparison of the
codes deployed in the Raspberry Pi. In Section \ref{sec:measurements}, we show
the measurements in the Raspberry Pi of the mentioned metrics. Final
conclusions and future work are reviewed in Section \ref{sec:conclusions}.

>>>>

Our work is organized as follows: Section \ref{sec:testbed} introduces
the testbed system, its setup, connectivity, files configuration and
compilation of the Kodo library. Section \ref{sec:schemes} defines the
coding schemes employed in our study. Later, in Section \ref{sec:metrics}
we describe the considered metrics for performance comparison of the
codes deployed in the \ac{Raspi}. In Section \ref{sec:measurements}, we show
the measurements in the \ac{Raspi} of the mentioned metrics. Final
conclusions and future work are reviewed in Section \ref{sec:conclusions}.

### 2016-06-04 metrics.tex
Update Raspi acronym

To complement our study, we review the energy consumption of the Raspberry
Pi since this platform is deployed at a large scale in scenarios where (i)
energy is constrained to the lifetime of the device battery and (ii) the
devices could be established in locations that are unavailable for
regular maintenance. Typical use cases of this type of scenarios are
sensor applications where devices are positioned for measurement retrieval
without any supervision for large periods fo time.

\subsection{Goodput}
We consider the goodput defined as the ratio of the useful delivered
information at the application layer and the total delivery time. We focus
on the goodput considering only the coding process, i.e. we assume that
the application data has been properly generated before encoding and
also correctly post-processed after decoding. In this way, we define
the goodput for either an encoder or a decoder as follows:
%
\begin{align} \label{eq:goodput}
Goodput = \frac{gB}{T_{proc}} ~ [\mathrm{Bytes/s}]
\end{align}
%
In \eqref{eq:goodput}, $B$ and $g$ are the packet and generation size
as defined previously and both represented the data to be processed. For
goodput measurements, we are concerned in quantifying the processing time
for either encoding or decoding $g$ \ac{l.i.} packets. Thus, $T_{proc}$ is
the processing time required for this processing. In the next subsections, we
define two time bechmarks available in [\textbf{Chres repos here}].
The purpose of the benchmarks is to quantify the processing time for
any of the coding scheme considered in Section \ref{sec:schemes}.

>>>>

To complement our study, we review the energy consumption of the \ac{Raspi}
since this platform is deployed at a large scale in scenarios where (i)
energy is constrained to the lifetime of the device battery and (ii) the
devices could be established in locations that are unavailable for
regular maintenance. Typical use cases of this type of scenarios are
sensor applications where devices are positioned for measurement retrieval
without any supervision for large periods fo time.

\subsection{Goodput}
We consider the goodput defined as the ratio of the useful delivered
information at the application layer and the total processing time. We focus
on the goodput considering only the coding process, i.e. we assume that
the application data has been properly generated before encoding and
also correctly post-processed after decoding. In this way, we define
the goodput for either an encoder or a decoder as follows:
%
\begin{align} \label{eq:goodput}
Goodput = \frac{gB}{T_{proc}} ~ [\mathrm{Bytes/s}]
\end{align}
%
In \eqref{eq:goodput}, $B$ and $g$ are the packet and generation size
as defined previously and both represented the data to be processed. For
goodput measurements, we are concerned in quantifying the processing time
for either encoding or decoding $g$ \ac{l.i.} packets. Thus, $T_{proc}$ is
the processing time required for this processing. In the next subsections, we
define two time bechmarks available in [\textbf{Chres repos here}].
The purpose of the benchmarks is to quantify the processing time for
any of the coding schemes considered in Section \ref{sec:schemes}.

### 2016-06-04 schemes.tex
Update Raspi acronym

In this section, we present a description of the considered coding schemes
measured with our the Raspberry Pi testbed. We introduce a definition
for the primitive coding operations, e.g. encoding, decoding and recoding
(where applies) for each coding scheme. Later, we address particular schemes
which are obtained by modifying the basic coding operations that provide
better processing speeds, particularly relevant for the Raspberry Pi.
Finally, we include a review of the

>>>>

In this section, we present a description of the considered coding schemes
measured with our the \ac{Raspi} testbed. We introduce a definition
for the primitive coding operations, e.g. encoding, decoding and recoding
(where applies) for each coding scheme. Later, we address particular schemes
which are obtained by modifying the basic coding operations that provide
better processing speeds, particularly relevant for the \ac{Raspi}.
Finally, we include a review of the

### 2016-06-04 testbed.tex
Update Raspi acronym

We proceed to define the testbed environment by each of its conforming
parts. We later indicate a procedure to set up the testbed. In this procedure,
we summarize all the details regarding on setting the connectivity and
system files for a set of Raspberry Pis in a centralized fashion.
Afterwards, we indicate how to cross-compile the Kodo library in an
easy way. Finally, we provide further information about Kodo itself in terms
of testing, other platforms supported and source code documentation
for further references.

@@

The Raspberry Pis are all equipped with an ethernet interface (eth0) and
connected by wire to the university's \ac{LAN}. The university network
administration was provided
the \ac{MAC} addresses of the ethernet interfaces to automatically assign
the devices the desired \ac{IP} addresses as shown in the figure. In a personal
\ac{LAN}, another option could be to manually provide the devices with static
\ac{IP}s.

@@

Finally, it should be possible to store data that are generated from
simulations on the memory cards that are in the individual Raspberry Pi.
%We using a overlaying filesystem that 
%We will illustrate
%how that can be done. In the case of temporary changes, it is important that
%all changes can be easily reverted. We put a \ac{RAM} filesystem on top of
%the the root filesystem to .
%the configure the
%devices to behave according to their test specifications. We
%will also present how that might be done.

@@

The Raspberry Pis are all equipped with an ethernet interface (eth0) and
connected by wire to the university's \ac{LAN}. The university network
administration was provided
the \ac{MAC} addresses of the ethernet interfaces to automatically assign
the devices the desired \ac{IP} addresses as shown in the figure. In a personal
\ac{LAN}, another option could be to manually provide the devices with static
\ac{IP}s.
%that are
%assigned the \ac{IP} addresses in the figure and connected wired to the
%university's \ac{LAN}.

We use standard usb power adapters to power the Raspberry Pis.

@@

Besides the previous description (\textbf{Include compiling Kodo from the
RasPi from the scratch}), the testbed administrator can compile Kodo in its
personal workstation and transfer the generated binaries directly to
a path in the Raspberry Pi. To achieve this, we get a toolchain that
contains the binaries for the \texttt{raspberry-gxx49-arm-g++} compiler
for the Raspberry Pi. Therefore, we strongly recommend any testbed
administrator to do the following procedure. In what follows, we provide
the instructions considering that the NFS server uses the \texttt{\$HOME}
directory as the working directory. However, the administrator may choose
some other working directory of its preference if desired.

@@

\item Download the Raspberry Pi toolchain for 64-bit Linux from: \\
\texttt{http://buildbot.steinwurf.dk/toolchains/linux/} to your
\texttt{\$HOME} directory. \\

@@

\item Navigate to the repository and configure \texttt{waf} by typing:
\texttt{python config.py} and select the 16th ``make specification'' file
for the Raspberry Pi, e.g. option \texttt{[16]cxx\_raspberry\_gxx49\_arm}
presented by the file.

This command configures \texttt{waf} to use the proper compiler and its
required flags to generate the binaries for the Raspberry Pi. If the
configuration was correct, the output will indicate:
\texttt{'configure' finished successfully (X.XXXs)}, where \texttt{X.XXX}
is total time in seconds for configuring the project in the server. \\

\item Execute \texttt{python waf build}. If the build process was
successful, the generated binaries for the Raspberry Pi should be located
in \texttt{build/cxx\_raspberry\_gxx49\_arm} in the Kodo repository.
\textbf{Indicate how the binary files should look like}.

Once this procedure is made, the testbed administrator can relocate the
generated binary files to Raspberry Pi's through the network as desired
by using the \texttt{scp} command during the configuration step.

@@

\subsection{Kodo Builds for the Raspberry Pi, Platform Support and Documentation}

You can check the build status of Kodo, Fifi and other relevant projects through
their respective repository master branch on our buildbot page
\cite{steinwurf2016buildbot}. Our buildbot displays the status of the builds for
Raspbian 8 and GCC 4.9 for the ARM architecture which is the relevant one for the
Raspberry Pi. At the link, you can check build status and build statistics. Also,
documentation about Kodo basics with a tutorial is available at \cite{kododocs}.

>>>>

We proceed to define the testbed environment by each of its conforming
parts. We later indicate a procedure to set up the testbed. In this procedure,
we summarize all the details regarding on setting the connectivity and
system files for a set of \ac{Raspi}s in a centralized fashion.
Afterwards, we indicate how to cross-compile the Kodo library in an
easy way. Finally, we provide further information about Kodo itself in terms
of testing, other platforms supported and source code documentation
for further references.

@@

The \ac{Raspi}s are all equipped with an ethernet interface (eth0) and
connected by wire to the university's \ac{LAN}. The university network
administration was provided
the \ac{MAC} addresses of the ethernet interfaces to automatically assign
the devices the desired \ac{IP} addresses as shown in the figure. In a personal
\ac{LAN}, another option could be to manually provide the devices with static
\ac{IP}s.

@@

Finally, it should be possible to store data that are generated from
simulations on the memory cards that are in the individual \ac{Raspi}.
%We using a overlaying filesystem that
%We will illustrate
%how that can be done. In the case of temporary changes, it is important that
%all changes can be easily reverted. We put a \ac{RAM} filesystem on top of
%the the root filesystem to .
%the configure the
%devices to behave according to their test specifications. We
%will also present how that might be done.

@@

The \ac{Raspi}s are all equipped with an ethernet interface (eth0) and
connected by wire to the university's \ac{LAN}. The university network
administration was provided
the \ac{MAC} addresses of the ethernet interfaces to automatically assign
the devices the desired \ac{IP} addresses as shown in the figure. In a personal
\ac{LAN}, another option could be to manually provide the devices with static
\ac{IP}s.
%that are
%assigned the \ac{IP} addresses in the figure and connected wired to the
%university's \ac{LAN}.

We use standard usb power adapters to power the \ac{Raspi}s.

@@

Besides the previous description (\textbf{Include compiling Kodo from the
RasPi from the scratch}), the testbed administrator can compile Kodo in its
personal workstation and transfer the generated binaries directly to
a path in the \ac{Raspi}. To achieve this, we get a toolchain that
contains the binaries for the \texttt{raspberry-gxx49-arm-g++} compiler
for the \ac{Raspi}. Therefore, we strongly recommend any testbed
administrator to do the following procedure. In what follows, we provide
the instructions considering that the NFS server uses the \texttt{\$HOME}
directory as the working directory. However, the administrator may choose
some other working directory of its preference if desired.

@@

\item Download the \ac{Raspi} toolchain for 64-bit Linux from: \\
\texttt{http://buildbot.steinwurf.dk/toolchains/linux/} to your
\texttt{\$HOME} directory. \\

@@

\item Navigate to the repository and configure \texttt{waf} by typing:
\texttt{python config.py} and select the 16th ``make specification'' file
for the \ac{Raspi}, e.g. option \texttt{[16]cxx\_raspberry\_gxx49\_arm}
presented by the file.

This command configures \texttt{waf} to use the proper compiler and its
required flags to generate the binaries for the \ac{Raspi}. If the
configuration was correct, the output will indicate:
\texttt{'configure' finished successfully (X.XXXs)}, where \texttt{X.XXX}
is total time in seconds for configuring the project in the server. \\

\item Execute \texttt{python waf build}. If the build process was
successful, the generated binaries for the \ac{Raspi} should be located
in \texttt{build/cxx\_raspberry\_gxx49\_arm} in the Kodo repository.
\textbf{Indicate how the binary files should look like}.

Once this procedure is made, the testbed administrator can relocate the
generated binary files to the \ac{Raspi}s through the network as desired
by using the \texttt{scp} command during the configuration step.

@@

\subsection{Kodo Builds for the \ac{Raspi}, Platform Support and Documentation}

You can check the build status of Kodo, Fifi and other relevant projects
through their respective repository master branch on our buildbot page
\cite{steinwurf2016buildbot}. Our buildbot displays the status of the builds
for Raspbian 8 and GCC 4.9 for the ARM architecture which is the relevant one
for the \ac{Raspi}. At the link, you can check build status and build
statistics. Also, documentation about Kodo basics with a tutorial is available
at \cite{kododocs}.
